---
title: "_Clipplings"
author: "Matthew Hamilton"
date: "18/02/2021"
output: html_document
---

## Matched groups details


## Single group details
The `r sngl_grp_ds_tb %>% dplyr::filter(round == "Baseline") %>% nrow()` baseline records are summarised below.
```{r}
(sngl_grp_ds_tb %>% dplyr::filter(round == "Baseline"))[4:7] %>% summary()
```
The summary of `r sngl_grp_ds_tb %>% dplyr::filter(round == "Follow-up") %>% nrow()` follow-up records:
```{r}
(sngl_grp_ds_tb %>% dplyr::filter(round == "Follow-up"))[4:7] %>% summary()
```
We add a cost variable with gamma distributed costs that have  mean values of \$300 for a defined period up to baseline and \$1500 for the costs accrued between baseline and follow-up. 
The resulting single group dataset contains `r nrow(sngl_grp_ds_tb)` records, with a snapshot provided below.

```{r}
sngl_grp_ds_tb  %>% head()
```

# Types of economic analysis facilitated by youthu
There are four types of economic analysis in youth mental health that youthu can help with:

- Extending cost-consequence and cost-effectiveness economic evaluations to include cost-utility analyses that are easier for healthcare policy-makers to interpret;
- Extending efficacy trials with a modelled analysis exploring the plausibility of the potential cost-effectiveness of study interventions;
- Extending single group datasets (e.g. health service records, pilot studys) with a modelled analysis of the potential for hypothethised interventions being cost-effective; and
- Assessing the potential economic value of alternative intervention research proposals.
```{r}
# var_nms_chr <- paste0(c("costs_dbl","qalys_dbl", paste0(c("PHQ9","SOFAS","AQOL6D_HU"), "_change_dbl")),"_Follow-up")
  
```

```{r}
# named_mat <- bootstraps_ls$t
# colnames(named_mat) <- names(bootstraps_ls$t0)
```

```{r}
evpi_ls <- BCEA::evppi(names(he_smry_ls$bootstraps_ls$t0)[c(1,3)],
            input = he_smry_ls$named_mat,
            he = he_smry_ls$ce_res_ls)
```
```{r fig.width=6}
BCEA::plot.evppi(evpi_ls, graph = "ggplot2")
```

```{r fig.width=6}
BCEA::eib.plot(he_smry_ls$ce_res_ls)
```

```{r fig.width=6}
BCEA::evi.plot(he_smry_ls$ce_res_ls,
               graph = "ggplot2")
```

```{r fig.width=6}
BCEA::ceac.plot(he_smry_ls$ce_res_ls,
                graph = "ggplot2",
          #title = "my title",
          line_colors = "green",
          theme = ggplot2::theme_dark())

```
