---
title: "Predict Quality Adjusted Life Years (QALYs)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Predict Quality Adjusted Life Years (QALYs)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  output_type_1L_chr: HTML
---

```{r message=FALSE}
library(youthu)
```
This vignette outlines a workflow for:

- Searching, selecting and retrieving transfer to utility models;
- Preparing a prediction dataset for use with a selected transfer to utility model; and
- Applying the selected transfer to utility model to a prediction dataset to predict Quality Adjusted Life Years (QALYs).

The practical value of implementing such a workflow is discussed in the [economic analysis vignette](Economic_Analysis.html) and [a scientific manuscript](https://www.medrxiv.org/content/10.1101/2021.07.07.21260129v2.full). Note, this example uses fake data - it should should not be used to inform decision making.

## Search, select and retrieve transfer to utility models
To identify datasets that contain transfer to utility models compatible with youthu (ie those developped with the [TTU package](https://ready4-dev.github.io/TTU/index.html)), you can use the `get_ttu_dv_dss` function. The function searches specified dataverses (in the below example, the [TTU dataverse](https://dataverse.harvard.edu/dataverse/TTU)) for datasets containing output from the TTU package.

```{r}
ttu_dv_dss_tb <- get_ttu_dv_dss("TTU")
```

The `ttu_dv_dss_tb` table summarises some pertinent details about each dataset containing TTU models found by the preceding command. These details include a link to any scientific summary (the "Article" column) associated with a dataset.

```{r ttudss, echo = F, tab.cap='Transfer to Utility Datasets', tab.id = 'ttudss', results="asis"}
ttu_dv_dss_tb[,c(1:3,5)] %>% 
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          use_lbls_as_col_nms_1L_lgl = T,
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```

To identify models that predict a specified type of health utility from one or more of a specified subset of predictors, use:

```{r }
mdls_lup <- get_mdls_lup(ttu_dv_dss_tb = ttu_dv_dss_tb,
                         utility_type_chr = "AQoL-6D",
                         mdl_predrs_in_ds_chr = c("PHQ9 total score",
                                                  "SOFAS total score"))
```

The preceding command will produce a lookup table with information that includes the catalogue names of models, the predictors used in each model and the analysis that generated each one.

```{r mdlslupsel, echo = F, tab.cap='Selected elements from Models Look-Up Table', tab.id = 'mdlslupsel', results="asis"}
mdls_lup[,c(1,2,5)] %>% 
  ready4use::add_labels_from_dictionary(dictionary_tb = tibble::tibble(var_nm_chr = names(mdls_lup)[c(1,2,5)],
                                                                       var_desc_chr = c("Catalogue reference", "Predictors", "Analysis"))) %>%
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          use_lbls_as_col_nms_1L_lgl = T,
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```

To review the summary information about the predictive performance of a specific model, use:

```{r}
get_dv_mdl_smrys(mdls_lup,
                 mdl_nms_chr = "PHQ9_SOFAS_1_OLS_CLL")

```

More information about a selected model can be found in the online model catalogue, the link to which can be obtained with the following command:

```{r results='asis'}
get_mdl_ctlg_url(mdls_lup,
                 mdl_nm_1L_chr = "PHQ9_SOFAS_1_OLS_CLL")
```

## Prepare a prediction dataset for use with a selected transfer to utility model

### Import data
You can now import and inspect the dataset you plan on using for prediction. In the below example we use fake data.

```{r }
data_tb <- make_fake_ds_one()
```
```{r dstb, echo = F, tab.cap='Illustrative example of a prediction dataset', tab.id = 'dstb', results="asis"}
data_tb %>% 
  head() %>%
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```

### Confirm dataset can be used as a prediction dataset
The prediction dataset must contain variables that correspond to all the predictors of the model you intend to apply. The allowable range and required class of each predictor variable are described in the `min_val_dbl`,	`max_val_dbl` and	`class_chr` columns of the model predictors lookup table, which can be accessed with a call to the `get_predictors_lup` function.

```{r}
predictors_lup <- get_predictors_lup(mdls_lup = mdls_lup,
                                     mdl_nm_1L_chr = "PHQ9_SOFAS_1_OLS_CLL")
```
```{r predluptb, echo = F, tab.cap='Model predictors lookup table', tab.id = 'predluptb', results="asis"}
predictors_lup %>% 
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```

The prediction dataset must also include both a unique client identifier variable and a measurement time-point identifier variable (which must be a `factor`{.R} with two levels). The dataset also needs to be in long format (ie where measures at different time-points for the same individual are stacked on top of each other in separate rows). We can confirm these conditions hold by creating a dataset metadata object using the `make_predn_metadata_ls` function. In creating the metadata object, the function checks that the dataset can be used in conjunction with the model specified at the `mdl_nm_1L_chr` argument. If the prediction dataset uses different variable names for the predictors to those specified in the `predictors_lup`{.R} lookup table, a named vector detailing the correspondence between the two sets of variable names needs to be passed to the `predr_vars_nms_chr` argument. Finally, if you wish to specify a preferred variable name to use for the predicted utility values when applying the model, you can do this by passing this name to the `utl_var_nm_1L_chr` argument.

```{r}
predn_ds_ls <- make_predn_metadata_ls(data_tb,
                                      id_var_nm_1L_chr = "UID",
                                      msrmnt_date_var_nm_1L_chr = "Date",
                                      predr_vars_nms_chr = c(PHQ9 = "PHQ_total",SOFAS = "SOFAS_total"),
                                      round_var_nm_1L_chr = "Timepoint",
                                      round_bl_val_1L_chr = "Baseline",
                                      utl_var_nm_1L_chr = "AQoL6D_HU",
                                      mdls_lup = mdls_lup,
                                      mdl_nm_1L_chr = "PHQ9_SOFAS_1_OLS_CLL")
```


## Apply the selected transfer to utility model to a prediction dataset to predict Quality Adjusted Life Years (QALYs)

### Predict health utility at baseline and follow-up timepoints
To generate utility predictions we use the `add_utl_predn`{.R} function. The function needs to be supplied with the prediction dataset (the value passed to argument `data_tb`) and the validated prediction metadata object we created in the previous step. 
```{r }
data_tb <- add_utl_predn(data_tb,
                         predn_ds_ls = predn_ds_ls)

```
By default the `add_utl_predn`{.R} function samples model parameter values based on a table of model coefficients when making predictions and constrains predictions to an allowed range. You can override these defaults by adding additional arguments `new_data_is_1L_chr = "Predicted"` (which uses mean parameter values), `force_min_max_1L_lgl = F` (removes range constraint) and (if the source dataset makes available downloadable model objects) `make_from_tbl_1L_lgl = F`. These settings will produce different predictions. It is strongly recommended that you consult the model catalogue (see above) to understand how such decisions may affect the validity of the predicted values that will be generated.


```{r updtb, echo = F, tab.cap='Prediction dataset with predicted utilities', tab.id = 'updtb', results="asis"}
data_tb %>% 
  head() %>%
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```

Our health utility predictions are now available for use and are summarised below.

```{r }
summary(data_tb$AQoL6D_HU)
```

### Calculate QALYs
The last step is to calculate Quality Adjusted Life Years, using a method assuming a linear rate of change between timepoints.

```{r}
data_tb <- data_tb %>% add_qalys_to_ds(predn_ds_ls = predn_ds_ls,
                                       include_predrs_1L_lgl = F,
                                       reshape_1L_lgl = F)
```

```{r qalytb, echo = F, tab.cap='Prediction dataset with QALYs', tab.id = 'qalytb', results="asis"}
data_tb %>% 
  head() %>%
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
    add_to_row_ls = NULL) 

```

```{r include=FALSE}
#Source: https://stackoverflow.com/questions/60042689/how-to-include-the-description-of-a-data-set-in-rmarkdown
help_text <- function(...) {
  file <- help(...)
  path <- dirname(file)
  dirpath <- dirname(path)
  pkgname <- basename(dirpath)
  RdDB <- file.path(path, pkgname)
  rd <- tools:::fetchRdDB(RdDB, basename(file))
  capture.output(tools::Rd2txt(rd, out="", options=list(underline_titles=FALSE)))
}
# help_text(add_aqol6d_predn_to_ds) %>%
#   purrr::map_chr(~.x %>% stringr::str_trim())
```
