---
title: "Predict Adolescent AQoL-6D"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Predict Adolescent AQoL-6D}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  output_type_1L_chr: HTML
---

```{r message=FALSE}
library(magrittr)
library(youthu)
```

## Determining if youthu models can be applied to a dataset
To use youthu to predict health utility values for records in a dataset, it is necessary for that dataset to include one or more of the predictors used in the youthu models. To verify that this is the case it is necessary to identify the predictors used in youthu models and to then compare these to the variables in the dataset being used for prediction.

### Identify measures that can be used to predict health utility with youthu models
To identify the predictors in youthu models consult the `predictors_lup`{.R} lookup table.

```{r results='hide'}
data("predictors_lup")
predictors_lup[,1:5]
```
```{r predrstb, echo = F, tab.cap='Predictors used in youthu models', tab.id = 'predrstb', results="asis"}
predictors_lup[,1:5] %>%
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```

### Match model predictors to the prediction dataset
Next, we can identify which of the above variables are present in the dataset we plan to use for prediction. To illustrate this step, we have created a prediction dataset that is a modified and simplified version of the study [replication dataset](https://ready4-dev.github.io/youthvars/articles/Replication_DS.html) distributed with the youthvars package. This illustrative dataset reflects the likely real-world circumstance in which the prediction dataset uses a different naming convention to that of the dataset we used to estimate the youth models. 

```{r }
data_tb <- youthvars::replication_popl_tb %>%
  dplyr::select(fkClientID,round,PHQ9, SOFAS) %>% #
  dplyr::arrange(fkClientID) %>%
  na.omit() %>%
  dplyr::rename(UID = fkClientID,
                Timepoint = round,
                PHQ_total = PHQ9,
                SOFAS_total = SOFAS) %>%
  dplyr::mutate(SOFAS_total = as.integer(round(SOFAS_total,0))) %>%
  tibble::as_tibble()
```
```{r results='hide'}
data_tb %>% head()
```
```{r dstb, echo = F, tab.cap='Illustrative example of a prediction dataset', tab.id = 'dstb', results="asis"}
data_tb %>% 
  head() %>%
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```

A visual inspection of the illustrative prediction dataset we have created reveals that the only predictors that match those included in youthu models are PHQ9 and SOFAS (the fifth and seventh rows of the `predictors_lup`{.R} lookup table). We can create a character vector to store this information as we will need it later. Note, the names stored in `mdl_predrs_in_ds_chr` must use the naming convention specified in the `predictors_lup`{.R} lookup table and not any different naming convention that may be used in the prediction dataset for the equivalent variables (in this case `PHQ_total` and `SOFAS_total`).

```{r}
mdl_predrs_in_ds_chr <- predictors_lup$short_name_chr[c(5,7)]
mdl_predrs_in_ds_chr
```
### Predictor data integrity checks
Another requirement is that the predictor variables in the prediction dataset must contain no missing values and conform to the requirements of class (in this case `integer`{.R}) and minimum and maximum values that are specified in the `predictors_lup`{.R} lookup table. We can confirm that these conditions hold by either visually inspecting the output of calls to base R `summary`{.R} and `class`{.R} functions or by using youthu's [data integrity tools](Data_Integrity.html).

#### Visual inspection checks
 
```{r}
summary(data_tb$PHQ_total)
```

```{r}
class(data_tb$PHQ_total)
```

```{r}
summary(data_tb$SOFAS_total)
```

```{r}
class(data_tb$SOFAS_total)
```

#### Checks using youthvars classes
We can pass the values of the `PHQ9_total` and `SOFAS_total` variables in our dataset to the functions for creating instances of the `youthvars_phq9` and `youthvars_sofas` classes from the [youthvars](https://ready4-dev.github.io/youthvars/) package. If no error message is thrown, then the data is consistent with the requirements for variable.

```{r}
phq9_test <- youthvars::youthvars_phq9(data_tb$PHQ_total) 
phq9_test %>% class()
```
```{r}
sofas_test <- youthvars::youthvars_sofas(data_tb$SOFAS_total) 
sofas_test %>% class()
```
### Confirm dataset meets other structural requirements
In addition to containing at least one match for the predictors used in youthu models, the prediction dataset must include both a unique client identifier variable (`character`{.R} class) and a measurement time-point identifier variable (`factor`{.R} class, with two levels). The dataset also needs to be in long format (ie where measures at different time-points for the same individual are stacked on top of each other in separate rows). We can confirm these conditions hold both by visual inspection of the dataset and investigation of the classes of the unique identified and measurement timepoint variables.

```{r}
data_tb$UID %>% class()
```
```{r}
data_tb$Timepoint %>% class()
levels(data_tb$Timepoint)
```

## Identify candidate models for available predictors
We can use the `get_mdls_using_predrs`{.R} function to identify which models can be used with the predictors present in the prediction dataset. The only argument that we need to pass to that function is our previously created `mdl_predrs_in_ds_chr`{.R} vector of predictors that are common to both our prediction dataset and youthu models. 

```{r message=FALSE}
data("mdls_lup", package ="youthu")
candidate_mdls_tb <- get_mdls_using_predrs(mdl_predrs_in_ds_chr)
```
```{r results='hide'}
candidate_mdls_tb
```
```{r candmdlstb, echo = F, tab.cap='youthu models that can be applied to the prediction dataset', tab.id = 'candmdlstb', results="asis"}
candidate_mdls_tb %>%
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL)
```

## Select and download preferred model
All youthu models are publicly available in a Dataverse repository (https://doi.org/10.7910/DVN/JC6PTV). The following function call will download the desired model from the data repository. To download a choosen model, use the `get_mdl_from_dv`{.R} function. In the below example we have selected the fourth model in the candidate models table (`r candidate_mdls_tb$mdl_nms_chr[4]`). We suggest consulting the scientific summary of the study that produced the youthu models to help guide your choice of which candidate model to select.

```{r warning=FALSE}
model_mdl <- get_mdl_from_dv(candidate_mdls_tb$mdl_nms_chr[4])
```

```{r}
model_mdl 
```
```{r include=F}
# TTU::make_brms_mdl_print_ls(mdl_ls = model_mdl,
#                                             label_stub_1L_chr = 5,
#                                             caption_1L_chr = "ANOTHER TEST",
#                                             output_type_1L_chr = "HTML")
```

## Predict utility
To generate utility predictions we use the `add_aqol6d_predn_to_ds`{.R} function. The function needs to be supplied with the prediction dataset (the value passed to argument `data_tb`), the model we are applying to that dataset (`model_mdl`), the transformation associated with that model (`tfmn_1L_chr`, which can be identified by passing the name of the selected model to the `get_tfmn_from_lup`{.R} function) and details of the variable names used in the prediction dataset for the unique identifier (`id_var_nm_1L_chr`) and measurement timepoint (`round_var_nm_1L_chr`) and, for the later, the factor level that represents the baseline timepoint (`round_bl_val_1L_chr`). Additionally, if the prediction dataset uses different variable names for the predictors to those specified in the `predictors_lup`{.R} lookup table, a named vector detailing the correspondence between the two sets of variable names needs to be passed to the `predr_vars_nms_chr` argument. Finally, if you wish to specify the desired variable name to use for the predicted utility values generated by the function, you can do this by passing this name to the `utl_var_nm_1L_chr` argument.

```{r}
updated_data_tb <- add_aqol6d_predn_to_ds(data_tb = data_tb,
                                   model_mdl = model_mdl,
                                   predr_vars_nms_chr = c(PHQ9 = "PHQ_total", SOFAS = "SOFAS_total"),
                                   tfmn_1L_chr = get_tfmn_from_lup(candidate_mdls_tb$mdl_nms_chr[4]),
                                   utl_var_nm_1L_chr = "AQoL6D_HU",
                                   id_var_nm_1L_chr = "UID",
                                   round_var_nm_1L_chr = "Timepoint",
                                   round_bl_val_1L_chr = "Baseline")

```
```{r results='hide'}
updated_data_tb %>% head()
```
```{r updtb, echo = F, tab.cap='Prediction dataset with predicted utilities', tab.id = 'updtb', results="asis"}
updated_data_tb %>% 
  head() %>%
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```

Our health utility predictions are now available for use and are summarised below.

```{r}
summary(updated_data_tb$AQoL6D_HU)
```


```{r include=FALSE}
#Source: https://stackoverflow.com/questions/60042689/how-to-include-the-description-of-a-data-set-in-rmarkdown
help_text <- function(...) {
  file <- help(...)
  path <- dirname(file)
  dirpath <- dirname(path)
  pkgname <- basename(dirpath)
  RdDB <- file.path(path, pkgname)
  rd <- tools:::fetchRdDB(RdDB, basename(file))
  capture.output(tools::Rd2txt(rd, out="", options=list(underline_titles=FALSE)))
}
help_text(add_aqol6d_predn_to_ds) %>%
  purrr::map_chr(~.x %>% stringr::str_trim())
```
