---
title: "Predict Adolescent AQoL-6D"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Predict Adolescent AQoL-6D}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  output_type_1L_chr: HTML
---

```{r message=FALSE}
library(magrittr)
# library(TTU)
library(youthu)
```

## Identifying suitable transfer to utility algorithms
To use youthu functions to predict health utility values for records in a dataset, it is necessary to identify transfer to utility models that use predictors that are also present in the prediction dataset. 

To identify datasets that contain transfer to utility models compatible with youthu (ie those developped with the [TTU package](https://ready4-dev.github.io/TTU/index.html)), you can use the following command. Note, the following example retrieves models created from fake data and is for instructional purposes only - these models should not be used to inform decision making.

```{r}
ttu_dv_dss_tb <- get_ttu_dv_dss("firstbounce")
```

```{r ttudss, echo = F, tab.cap='Transfer to Utility Datasets', tab.id = 'ttudss', results="asis"}
ttu_dv_dss_tb %>% 
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          use_lbls_as_col_nms_1L_lgl = T,
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```

To identify models that predict a specified type of health utility from one or more of a specified subset of predictors, use:

```{r}
mdls_lup <- get_mdls_lup(ttu_dv_dss_tb = ttu_dv_dss_tb,
                         utility_type_chr = "AQoL-6D",
                         mdl_predrs_in_ds_chr = c("PHQ9 total score",
                                                  "SOFAS total score"))
```

The preceding command will produce a lookup table with information that includes the catalogue names of models, the predictors used in each model and the analysis that generated each one.

```{r}
mdls_lup[,c(1,2,5)]
```

To review the summary information about the predictive performance of a specific model, use:

```{r}
get_dv_mdl_smrys(mdls_lup,
                 mdl_nms_chr = "SOFAS_1_OLS_CLL")

```

More information about a selected model can be found in the online model catalogue, the link to which can be obtained with the following command:

```{r}
get_mdl_ctlg_url(mdls_lup,
                 mdl_nm_1L_chr = "SOFAS_1_OLS_CLL")
```

## Prepare prediction dataset
First import and inspect the dataset you will be using for prediction. In the below example we use fake data.

```{r }
data_tb <- make_fake_ds_one()
```
```{r dstb, echo = F, tab.cap='Illustrative example of a prediction dataset', tab.id = 'dstb', results="asis"}
data_tb %>% 
  head() %>%
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```

### Confirm dataset can be used as a prediction dataset
The prediction dataset must contain variables that correspond to all the predictors of the model you intend to apply. The values of each predictor variable must be within the allowable range and be of a compatible class to the corresponding model predictor metadata.

```{r}
get_predictors_lup(mdls_lup = mdls_lup,
                   mdl_nm_1L_chr = "SOFAS_1_OLS_CLL")
```
The prediction dataset must include both a unique client identifier variable and a measurement time-point identifier variable (`factor`{.R} class, with two levels). The dataset also needs to be in long format (ie where measures at different time-points for the same individual are stacked on top of each other in separate rows). We can confirm these conditions hold by creating a dataset metadata object using the `make_valid_predn_ds_ls` function. In creating the metadata object, the function checks that the dataset can be used in conjunction with the model specified at the `mdl_nm_1L_chr` argument. If the prediction dataset uses different variable names for the predictors to those specified in the `predictors_lup`{.R} lookup table, a named vector detailing the correspondence between the two sets of variable names needs to be passed to the `predr_vars_nms_chr` argument. Finally, if you wish to specify a preferred variable name to use for the predicted utility values when applying the model, you can do this by passing this name to the `utl_var_nm_1L_chr` argument.

```{r}
predn_ds_ls <- make_predn_metadata_ls(data_tb,
                                      id_var_nm_1L_chr = "UID",
                                      predr_vars_nms_chr = c(SOFAS = "SOFAS_total"),
                                      round_var_nm_1L_chr = "Timepoint",
                                      round_bl_val_1L_chr = "Baseline",
                                      utl_var_nm_1L_chr = "AQoL6D_HU",
                                      mdls_lup = mdls_lup,
                                      mdl_nm_1L_chr = "SOFAS_1_OLS_CLL")
```


## Predict utility

To generate utility predictions we use the `add_utl_predn`{.R} function. The function needs to be supplied with the prediction dataset (the value passed to argument `data_tb`) and the validated prediction metadata object we created in the previous step. By default the function chooses a constrained simulation approach, based on a table of model coefficients when making predictions. You can override this by adding additional arguments `new_data_is_1L_chr = "Predicted"`, `force_min_max_1L_lgl = T` and (if the source dataset makes available downloadable models) `make_from_tbl_1L_lgl = F`. However, before doing so it is strongly recommended that you consult the model catalogue (see above) to understand how such a decision may affect the validity of the predicted values that will be generated.

```{r eval = T}
updated_data_tb <- add_utl_predn(data_tb,
                                 predn_ds_ls = predn_ds_ls)

```
```{r updtb, echo = F, tab.cap='Prediction dataset with predicted utilities', tab.id = 'updtb', results="asis"}
updated_data_tb %>% 
  head() %>%
  ready4show::print_table(output_type_1L_chr = params$output_type_1L_chr,
                          caption_1L_chr = knitr::opts_current$get("tab.cap"),
                          mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                          add_to_row_ls = NULL) 
```
Our health utility predictions are now available for use and are summarised below.

```{r}
summary(updated_data_tb$AQoL6D_HU)
```


```{r include=FALSE}
#Source: https://stackoverflow.com/questions/60042689/how-to-include-the-description-of-a-data-set-in-rmarkdown
help_text <- function(...) {
  file <- help(...)
  path <- dirname(file)
  dirpath <- dirname(path)
  pkgname <- basename(dirpath)
  RdDB <- file.path(path, pkgname)
  rd <- tools:::fetchRdDB(RdDB, basename(file))
  capture.output(tools::Rd2txt(rd, out="", options=list(underline_titles=FALSE)))
}
help_text(add_aqol6d_predn_to_ds) %>%
  purrr::map_chr(~.x %>% stringr::str_trim())
```
