---
title: "Predict health utility from psychological measures (PHQ-9 and GAD-7)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Predict health utility from psychological measures (PHQ-9 and GAD-7)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**This vignette article is abridged and modified version of [another article on predicting Quality Adjusted Life Years with youthu](https://ready4-dev.github.io/youthu/articles/Prediction_With_Mdls.html).**

## Motivation
This article illustrates how to make QALY predictions using a dataset in wide format with no health-utility measures but containing two psychological measures (GAD-7 and PHQ-9).

## Install youthu
If not already installed it will be necessary to install the `youthu` R library. As `youthu` is not yet available on CRAN, it will be necessary to install it directly from its GitHub repository using an R package like `remotes` or `devtools`.

```{r eval=FALSE}
# Uncomment and run if installation is required.
# utils::install.packages("devtools") 
# devtools::install_github("ready4-dev/youthu")
```

## Load required packages
We now load the libraries we will be using in subsequent steps. Note, both the `ready4`, `ready4show` and `ready4use` [ready4 framework](https://www.ready4-dev.com/) libraries will have been installed automatically when `youthu` was installed. The `specific` readyforwhatsnext module library and `dplyr`, `purrr`, `stringr` and `tidyr` CRAN libraries will have been installed at the same time.

```{r message=FALSE}
library(ready4)
library(ready4show)
library(ready4use)
library(specific)
library(youthu)
```

## Specify data sources
We begin by specifying the sources for our data. In this example, our data sources are online repositories.

```{r}
X <- Ready4useRepos(dv_nm_1L_chr = "fakes", dv_ds_nm_1L_chr = "https://doi.org/10.7910/DVN/HJXYKQ", 
                    dv_server_1L_chr = "dataverse.harvard.edu",
                    gh_repo_1L_chr = "ready4-dev/youthu", gh_tag_1L_chr = "v0.0.0.91125")
```

## Inspect dataset
We can now inspect the dataset we will be using to make predictions. As this is a demonstration article we are going to create a custom synthetic dataset. Our first step in doing so is to ingest a preexisting synthetic dataset (in long format) using the method explained in [another vignette article](https://ready4-dev.github.io/ready4use/articles/V_03.html)

```{r}
data_tb <- ingest(X, fls_to_ingest_chr = c("ymh_phq_gad_tb"), metadata_1L_lgl = F)
```

```{r echo=FALSE, eval = FALSE}
# Alternatively
# data_tb <- ingest(X, fls_to_ingest_chr = c("ymh_clinical_tb"), metadata_1L_lgl = F)
# data_tb <- data_tb %>% dplyr:: select(c("fkClientID", "round", 
#                                         "d_interview_date", 
#                                         "gad7_total", "phq9_total")) %>%
#   tidyr::pivot_wider(names_from = c("round"), 
#                      values_from = c("d_interview_date", "gad7_total", "phq9_total")) %>%
#   dplyr::rename_with(~stringr::str_replace(.x,"_Baseline","_t1") %>% 
#                        stringr::str_replace("_Follow-up","_t2") %>% 
#                        stringr::str_replace("_total",""))
```

Our resulting dataset has unique IDs for each participant (character class), timestamps for each data collection timepoint (Date class variables) and GAD-7 and PHQ-9 scores for each timepoint (integer class).


```{r }
data_tb %>% head() %>% ready4show::print_table(caption_1L_chr = "Dataset", output_type_1L_chr = "HTML") 
```

## Get mapping models
We retrieve details of relevant AQoL-6D mapping models for wither of the predictors we plan on using. How these models were derived is described in a [pre-print](https://doi.org/10.1101/2021.07.07.21260129) and details of model performance is included in catalogues available in an [open access data repository](https://doi.org/10.7910/DVN/DKDIB0).

```{r}
mdls_lup <- get_mdls_lup(ttu_dv_dss_tb = get_ttu_dv_dss("TTU"),
                         utility_type_chr = "AQoL-6D",
                         mdl_predrs_in_ds_chr = c("GAD7 total score", "PHQ9 total score"))
```

```{r }
mdls_lup[,c(1,2,5)] %>% 
  ready4show::print_table(caption_1L_chr = "Available models", output_type_1L_chr = "HTML") 
```

We select our preferred model and retrieve summary data about the model's predictor variables.

```{r }
predictors_lup <- get_predictors_lup(mdls_lup = mdls_lup, mdl_nm_1L_chr = "GAD7_PHQ9_1_OLS_CLL")
```

```{r }
exhibit(predictors_lup)
```

## Transform prediction dataset
To be used with the mapping models available to us, our prediction dataset needs to be in long format. We perform the necessary transformation.

```{r}
data_tb <- transform_ds_to_long(data_tb, predictors_chr = c("gad7", "phq9"),
                             msrmnt_date_var_nm_1L_chr = "d_interview_date", round_var_nm_1L_chr = "When")
```
We drop records where we are missing data for either GAD7 or PHQ9 at either timepoint.

```{r}
data_tb <- transform_ds_to_drop_msng(data_tb, predictors_chr = c("gad7", "phq9"), 
                                      uid_var_nm_1L_chr = "fkClientID")
```


We now predict AQoL-6D health utility for each case with complete data.

```{r}
predn_ds_ls <- make_predn_metadata_ls(data_tb,
                                      id_var_nm_1L_chr = "fkClientID",
                                      msrmnt_date_var_nm_1L_chr = "d_interview_date",
                                      predr_vars_nms_chr = c(GAD7 = "gad7", PHQ9 = "phq9"),
                                      round_var_nm_1L_chr = "When",
                                      round_bl_val_1L_chr = "t1",
                                      utl_var_nm_1L_chr = "AQoL6D_HU",
                                      mdls_lup = mdls_lup,
                                      mdl_nm_1L_chr = "GAD7_PHQ9_1_OLS_CLL")
data_tb <- add_utl_predn(data_tb, new_data_is_1L_chr = "Predicted", predn_ds_ls = predn_ds_ls)
```

Finally, we derive QALY predictions from the health utility measures at both time-points.

```{r}
data_tb <- data_tb %>% add_qalys_to_ds(predn_ds_ls = predn_ds_ls, include_predrs_1L_lgl = F, reshape_1L_lgl = T)
```

```{r }
data_tb %>% head() %>%
  ready4show::print_table(caption_1L_chr = "Final dataset", output_type_1L_chr = "HTML",
                          scroll_box_args_ls = list(width = "100%"))
```
